Prediction of PIK3CA mutation with gene expression

### 2020-04-22 plosOne/plosOne.Rmd
pdf
rpackage citations

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr rpackage. Gene expression data was get from GDAC firehose using RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included. Data of invasive ductal carcinomas were used for training set and data of invasive lobular carcinoma were used for test set. Number of observations were 530 in training set and 188 in test set.

## Selecting variable for modeling

@@

## Preprocessing

Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.

Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).

@@

The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of trainset was 0.74 and the testset AUPR was 0.771.

```{r}
load("../result.RData")
```

## performance of each cancer type

@@

PIK3CA prediction model performed better than RAS activation prediction model in terms of AUROC. The overall AUPR was

>>>>

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was get from GDAC firehose using RTCGAToolbox R package. [@samurRTCGAToolboxNewTool2014] ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included. Data of invasive ductal carcinomas were used for training set and data of invasive lobular carcinoma were used for test set. Number of observations were 530 in training set and 188 in test set.

@@

Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020]

Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR). The modelin process was done with tidymodels rpackage.  [@kuhnTidymodelsEasilyInstall2020]

@@

The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was 0.39. (Figure 1A)

## performance of each cancer type  
Because tht prevalence of PIK3CA mutation is vary across the cancer type, the performance of each cancer type was investigated. The AUROC and AUPR were positively correlated between train set and test set in cancer type subanalysis. (Figure 1B) The AUPR was high in cancer type with high PIK3CA mutation rate such as colon, brest, Uterus cancer types. The AUROC did not correlated with PIK3CA mutation rate of each cancer types. (Figure 1C)

@@

PIK3CA prediction model performed better than RAS activation prediction model in terms of AUROC. AUPR and the PIK3CA mutation rate interpretation. @davis_2006_relationship_proc23rdintconfmachlearn

### 2020-04-22 plosOne/plosOne.tex
rpackage citations

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
rpackage. Gene expression data was get from GDAC firehose using
RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative
and/or SISH negative breast cancers are included. Data of invasive
ductal carcinomas were used for training set and data of invasive
lobular carcinoma were used for test set. Number of observations were
530 in training set and 188 in test set.

\hypertarget{selecting-variable-for-modeling}{%
\subsection{Selecting variable for
modeling}\label{selecting-variable-for-modeling}}

@@

\hypertarget{preprocessing}{%
\subsection{Preprocessing}\label{preprocessing}}

Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.

Penalized logistic regression was applied to prediction modeling.
10-fold cross-validation with targe variable stratification was done
over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Model performance was evealuated
with the area under receiver operating characteristic (ROC) and area
under the ROC (AUROC) and precision recall (PR) and area under curve
(AUPR).

@@

The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of
trainset was 0.74 and the testset AUPR was 0.771.

@@

@@

PIK3CA prediction model performed better than RAS activation prediction
model in terms of AUROC. AUPR and the PIK3CA mutation rate
interpretation. {[}3{]}

@@

S1 Appendix. S2 Figure. S1 Table.\\
S2 Table.\\
S3 Table.

@@

\leavevmode\hypertarget{ref-davis_2006_relationship_proc23rdintconfmachlearn}{}%
3. Davis J, Goadrich M. The relationship between Precision-Recall and
ROC curves. Proceedings of the 23rd international conference on Machine
learning. Pittsburgh, Pennsylvania, USA: Association for Computing
Machinery; 2006. pp. 233--240.
doi:\href{https://doi.org/10.1145/1143844.1143874}{10.1145/1143844.1143874}

>>>>

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
rpackage.{[}3{]} Gene expression data was get from GDAC firehose using
RTCGAToolbox R package. {[}4{]} ER immunostain postive HER2 immunostain
negative and/or SISH negative breast cancers are included. Data of
invasive ductal carcinomas were used for training set and data of
invasive lobular carcinoma were used for test set. Number of
observations were 530 in training set and 188 in test set.

@@

Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.
{[}5{]}

Penalized logistic regression was applied to prediction modeling.
10-fold cross-validation with targe variable stratification was done
over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Model performance was evealuated
with the area under receiver operating characteristic (ROC) and area
under the ROC (AUROC) and precision recall (PR) and area under curve
(AUPR). The modelin process was done with tidymodels rpackage. {[}6{]}

@@

The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of
trainset was 0.66 and the testset AUPR was 0.39. (Figure 1A)

@@

Because tht prevalence of PIK3CA mutation is vary across the cancer
type, the performance of each cancer type was investigated. The AUROC
and AUPR were positively correlated between train set and test set in
cancer type subanalysis. (Figure 1B) The AUPR was high in cancer type
with high PIK3CA mutation rate such as colon, brest, Uterus cancer
types. The AUROC did not correlated with PIK3CA mutation rate of each
cancer types. (Figure 1C)

@@

PIK3CA prediction model performed better than RAS activation prediction
model in terms of AUROC. AUPR and the PIK3CA mutation rate
interpretation. {[}7{]}

@@

S1 Appendix.\\
S2 Figure.\\
S1 Table.\\
S2 Table.\\
S3 Table.

@@

\leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
3. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
genomics data server (CGDS). 2019.

\leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
4. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
PLoS One. 2014;9(9):e106397.

\leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
5. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
matrices. 2020.

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
6. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-davis_2006_relationship_proc23rdintconfmachlearn}{}%
7. Davis J, Goadrich M. The relationship between Precision-Recall and
ROC curves. Proceedings of the 23rd international conference on Machine
learning. Pittsburgh, Pennsylvania, USA: Association for Computing
Machinery; 2006. pp. 233--240.
doi:\href{https://doi.org/10.1145/1143844.1143874}{10.1145/1143844.1143874}
